The Algorand Virtual Machine

### 2021-03-07 tex/avm.tex
Fix minor readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

    If we consider a proper vertex coloring of \(G\), every color class will give us an independent set of
    transactions which can be validated concurrently. To achieve the highest parallelization, we need to color \(G\)
    with minimum number of colors. Thus, the chromatic number of the memory dependency graph shows how good a
    transaction set could be run concurrently.

    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
    solution. An approximate greedy algorithm will perform well enough in most circumstances. The block proposer is
    responsible for solving the graph coloring problem anda proposed block must determine the independent sets of
    transactions which can be run in parallel safely. Since with better parallelization a block can contain more
    transactions, a proposer is incentivized enough to find a good graph coloring.

@@

    To mitigate this problem, for calculating a user's stake at the time step \(t\), instead of using the raw ALGO
    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
    ALGO balance:
    \[
        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
    \]
    The agreement protocol, at the time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
    exponential moving average of the ALGO balance of the user \(u\) at the time step \(t\).

@@

    For calculating the exponential moving average of a user's balance at the time step \(t\), we can use the following
    recursive formula:
    \[
        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
    \]
    Where:
    \begin{itemize}
        \item The coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
        degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
        \item \(B_{u,t}\) is the balance of the user \(u\) at the time step \(t\).
        \item \(M_{u,t}\) is the EMA for the user \(u\) at the time step \(t\).
    \end{itemize}

@@

    \[
        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
    \]
    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
    we can obtain \(m\) from the following formula, when \(\alpha\) is small:
    \[
        m = \frac{1}{\alpha} \ln \left(\frac{\gamma - 1}{\lambda - 1}\right)
    \]

>>>>

    If we consider a proper vertex coloring of \(G\), every color class will give us an independent set of
    transactions which can be validated concurrently. To achieve the highest parallelization, we need to color \(G\)
    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
    transaction set could be run concurrently.

    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
    solution. An approximate greedy algorithm will perform well enough in most circumstances. The block proposer is the
    one who is responsible for solving the graph coloring problem and a proposed block must determine the independent
    sets of transactions which can be run in parallel safely. Since with better parallelization a block can contain more
    transactions, a proposer is incentivized enough to find a good graph coloring.

@@

    To mitigate this problem, for calculating a user's stake at the time step \(t\), instead of using the raw ALGO
    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
    ALGO balance:
    \[
        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
    \]
    Where:
    \begin{itemize}
        \item \(S_{u,t}\) is the stake of the user \(u\) at the time step \(t\).
        \item \(B_{u,t}\) is the ALGO balance of the user \(u\) at the time step \(t\).
        \item \(Trust_{u,t}\) is an estimated trust value for the user \(u\) at the time step \(t\).
    \end{itemize}

    The agreement protocol, at the time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
    exponential moving average of the ALGO balance of the user \(u\) at the time step \(t\).

@@

    For calculating the exponential moving average of a user's balance at the time step \(t\), we can use the following
    recursive formula:
    \[
        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
    \]
    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
    degree of
    weighting decrease, A higher \(\alpha\) discounts older observations faster.

@@

    \[
        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
    \]
    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
    \[
        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
    \]
